---
title: "midterm (parallel)"
author: "Lumi Huang"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Assume that your restaurant opens at 8am and closes at 5pm. Customers arrive at a rate of 4/hr. Between the hours of noon and 1 that rate increases to 6/hr. Suppose you have 3 tellers. We have studied 3 types of queuing systems**
**Given three tellers which queuing system has the shortest expected waiting time for the customers?**
Based on the results below. 1 line with parallel servers has the shortest expected customer waiting time.

**(1) 1 line with parallel servers**
```{r}
lambda.t <- function(time){		# the banks open from 8am to 5pm, so we set 	
	if(time >= 4 && time <= 5){	# the time from 0 to 9 and the rate is 4/hr 
		lambda <- 6		# but increase to 6 between 4 and 5.
	} else { lambda <- 4 }
	list(lamt=lambda)
}


#Generate Tt which will be assigned to ta: next arrival time

generate.Tt <- function(ta, t) {
	Tt <- t- (1/lambda.t(t)$lamt)*log(runif(1))	# next arrival time

	if (ta < 4) {		# ta is less than 4
			
		ta <- Tt
		if (ta > 4) { ta <- 4+((ta-4)*4/6) }	#if greater than 4, change rate	
	
	} else if ( 4 <= ta && ta <= 5 ) {	# 4 <= ta <= 5
		ta <- Tt
		if (ta > 5) { ta <- 5+((ta-5)*6/4) }	#if greater than 5 , change rate
	} else { 
		ta <- Tt
	}
	if ( ta > 9){
	  ta <- 9999	# To insure that nobody can enter the system
	}
	return(ta)
}	

N <- 1
cust.wait <- numeric(N)
Tp <- numeric(N)
depart <- numeric(N)
Timepast <- numeric(N)
i <- 0

for(i in 1:N){
  t <- 0 #time starts at 0
  T3 <- 9 #store closed
  A <- numeric() #arrival time of customer
  D <- numeric() #departure time of customer
  Na <- 0 #number of arrival by time t
  Nd <- 0
  u1 <- runif(1)
  Ta <- 0
  Ta <- generate.Tt(Ta, t) #generate the first arrival time
  n <- 0 #number of customer in the store
  i1 <- 0 #index of the customer at server 1 (i1 = 0 if no one at server 1)
  i2 <- 0 #index of the customer at server 2
  i3 <- 0 #index of the customer at server 3
  t1 <- 9999 #service completion time of customer being serve at server 1
  t2 <- 9999 #service completion time of customer being serve at server 2
  t3 <- 9999 #service completion time of customer being serve at server 3
  c1 <- 0 #number of customer serviced by casher 1 by time t
  c2 <- 0
  c3 <- 0
  #store open
  while(t < T3){
    #case 1: someone arrive
    if(Ta == min(Ta, t1, t2, t3) & Ta < T3){
      t <- Ta
      Na <- Na + 1
      A[Na] <- t #record the Na-th arrival time
      Ta <- generate.Tt(Ta, t) #generate the next arrival time
      if(n == 0 & i1 == 0 & i2 == 0 & i3 == 0){
        n <- 1
        #if the store is empty
        i1 <- Na #force this person to go to server 1
        Y1 <- rgamma(1, 1, 10) #service period
        t1 <- t + Y1
      }else if(n == 1 & i1 > 0 & i2 == 0 & i3 == 0){
        #if someone occupies server 1 & server 2 and 3 are empty
        n <- 2
        i2 <- Na #force this person to go to server 2
        Y2 <- rgamma(1, 1, 10)
        t2 <- t + Y2
      }else if(n == 1 & i1 == 0 & i2 > 0 & i3 == 0){
        n <- 2
        #if someone occupies server 2 & server 1 and 3 are empty
        i3 <- Na #force this person to go to server 3
        Y3 <- rgamma(1, 1, 10)
        t3 <- t + Y3
      }else if(n == 1 & i1 == 0 & i2 == 0 & i3 > 0){
        #if someone occupies server 3 & server 1 and 2 are empty
        n <- 2
        i1 <- Na #force this person to go to server 1
        Y1 <- rgamma(1, 1, 10)
        t1 <- t + Y1
      }else if(n == 2 & i1 > 0 & i2 > 0 & i3 == 0){
          #if someone occupies server 1 and 2 & server 3 is empty
        n <- 3
        i3 <- Na #force this person to go to server 3
        Y3 <- rgamma(1, 1, 10)
        t3 <- t + Y3
      }else if(n == 2 & i1 > 0 & i2 == 0 & i3 > 0){
        #if someone occupies server 1 and 3 & server 2 is empty
        n <- 3
        i2 <- Na
        Y2 <- rgamma(1, 1, 10)
        t2 <- t + Y2
      }else if(n == 2 & i1 == 0 & i2 > 0 & i3 > 0){
        #if someone occupies server 2 and 3 & server 1 is empty
        n <- 3
        i1 <- Na
        Y1 <- rgamma(1, 1, 10)
        t1 <- t + Y1
      }else if(n > 2){
        n <- n + 1
      }
    }
    else if(t1 == min(Ta, t1, t2, t3) & t1 < T3){
      #case 2: someone leaves server 1
      t <- t1
      c1 <- c1 + 1
      Nd <- Nd + 1
      D[i1] <- t
      if(n <= 3){
        n <- n - 1
        i1 <- 0
        t1 <- 9999
      }else if(n > 3){
        n <- n - 1
        m <- max(i1, i2, i3)
        i1 <- m + 1
        Y1 <- rgamma(1, 1, 10)
        t1 <- t + Y1
      }
    }else if(t2 == min(Ta, t1, t2, t3) & t2 < T3){
      #case 3: someone leaves server 2
      t <- t2
      c2 <- c2 + 1
      Nd <- Nd + 1
      D[i2] <- t
      if(n <= 3){
        n <- n - 1
        i2 <- 0
        t2 <- 9999
      }else if(n > 3){
        n <- n - 1
        m <- max(i1, i2, i3)
        i2 <- m + 1
        Y2 <- rgamma(1, 1, 10)
        t2 <- t + Y2
      }
    }else if(t3 == min(Ta, t1, t2, t3) & t3 < T3){
      #case 4: someone leaves server 3
      t <- t3
      c3 <- c3 + 1
      Nd <- Nd + 1
      D[i3] <- t
      if(n <= 3){
        n <- n - 1
        i3 <- 0
        t3 <- 9999
      }else if(n > 3){
        n <- n - 1
        m <- max(i1, i2, i3)
        i3 <- m + 1
        Y3 <- rgamma(1, 1, 10)
        t3 <- t + Y3
      }
    }
    if(min(Ta, t1, t2, t3) >= T3) break
  }
  #store is closed
  if(min(Ta, t1, t2, t3) >= T3 & n > 0){
    while(n > 0){
      #leave server 1
      if(t1 == min(t1, t2, t3)){
        t <- t1
        Nd <- Nd + 1
        c1 <- c1 + 1
        D[i1] <- t
        if(n <= 3 ){
          n <- n - 1
          i1 <- 0
          t1 <- 9999
        }else if(n > 3){
          n <- n - 1
          m <- max(i1, i2, i3)
          i1 <- m + 1
          Y1 <- rgamma(1, 1, 10)
          t1 <- t + Y1
        }
      }
      else if(t2 == min(t1, t2, t3)){
        #leaves server 2
        t <- t2
        Nd <- Nd + 1
        c2 <- c2 + 1
        D[i2] <- t
        if(n <= 3){
          n <- n - 1
          i2 <- 0
          t2 <- 9999
        }else if(n > 3){
          n <- n - 1
          m <- max(i1, i2, i3)
          i2 <- m + 1
          Y2 <- rgamma(1, 1, 10)
          t2 <- t + Y2
        }
      }
      else if(t3 == min(t1, t2, t3)){
        #leaves server 3
        t <- t3
        Nd <- Nd + 1
        c3 <- c3 + 1
        D[i3] <- t
        if(n <= 3){
          n <- n - 1
          i3 <- 0
          t3 <- 9999
        }else if(n >3){
          n <- n - 1
          m <- max(i1, i2, i3)
          i3 <- m + 1
          Y3 <- rgamma(1, 1, 10)
          t3 <- t + Y3
        }
      }
    }
  }
  cust.wait[i] <- mean(D - A)
  i <- i + 1
}
mean(cust.wait)
```


**(2) tandem (in n out)**
```{r}
lambda.t <- function(time){		# the banks open from 8am to 5pm, so we set 	
  if(time >= 4 && time <= 5){	# the time from 0 to 9 and the rate is 4/hr 
    lambda <- 6		# but increase to 6 between 4 and 5.
  } else { lambda <- 4 }
  list(lamt=lambda)
}


### subroutine ###
### Generate Tt which will be assigned to ta: next arrival time ###

generate.Tt <- function(ta, t) {
  Tt <- t- (1/lambda.t(t)$lamt)*log(runif(1))	# next arrival time
  
  if (ta < 4) {		# ta is less than 4
    
    ta <- Tt
    if (ta > 4) { ta <- 4+((ta-4)*4/6) }	#if greater than 4, change rate	
    
  } else if ( 4 <= ta && ta <= 5 ) {	# 4 <= ta <= 5
    ta <- Tt
    if (ta > 5) { ta <- 5+((ta-5)*6/4) }	#if greater than 5 , change rate
  } else { 
    ta <- Tt
  }
  if ( ta > 9){
    ta <- 9999	# To insure that nobody can enter the system
  }
  return(ta)
}	

N <- 100
cust.wait <- numeric(N)
i <- 1
for (i in 1:N){
  t <- 0 #time starts at 0
  T3 <- 9
  A <- numeric() #arrival time of customer at server 1
  D1 <- numeric() #departure time of customer at server 1
  D2 <- numeric() #departure time of customer at server 2
  D3 <- numeric() #departure time of customer at server 3
  Na <- 0 #number of arrival by time t
  Nd <- 0 #number of departure by time t
  u1 <- runif(1)
  Ta <- 0
  Ta <- generate.Tt(Ta, t) 
  n1 <- 0 #number of customers in line at server 1
  n2 <- 0 #number of customers in line at server 2
  n3 <- 0 #number of customers in line at server 3
  t1 <- 9999 #service completion time of customer being serve at server 1
  t2 <- 9999 #service completion time of customer being serve at server 2
  t3 <- 9999 #service completion time of customer being serve at server 3
  #store is open
  while(t < T3){
    #case 1: [store open] someone arrives at server 1
    if(Ta == min(Ta, t1, t2, t3) & Ta < T3){
      t <- Ta
      Na <- Na + 1
      n1 <- n1 + 1
      A[Na] <- t #arrival time at server 1 for customer i
      #generate Tt and reset Ta = Tt
      Ta <- generate.Tt(Ta, t) #generate the next arrival time
      if (n1 == 1){
        Y1 <- rgamma(1,1,10)
        t1 <- t + Y1		#t1=t+Y because system was empty and t1=infty
      }
    }else if(t1 == min(t1, t2, t3, Ta) & t1 <= T3){
      #case 2: [store open] someone leaves server 1
      t <- t1
      n1 <- n1 - 1
      n2 <- n2 + 1
      D1[Na - n1] <- t
      if(n1 == 0){
        t1 <- 9999
      }else{
        Y1 <- rgamma(1, 1, 10)
        t1 <- t + Y1
      }
      if(n2 == 1){
        Y2 <- rgamma(1, 1, 10)
        t2 <- t + Y2
      }
    }
    #case 3: [store open] someone leaves server 2 before someone arrives at server 2 
    #and before someone arrives at server 1
    else if(t2 == min(t1, t2, t3, Ta) & t2 <= T3){
      t <- t2
      n2 <- n2 - 1
      n3 <- n3 + 1
      D2[Na - n1 - n2] <- t
      if(n2 == 0){
        t2 <- 9999
      }
      else{
        Y2 <- rgamma(1, 1, 10)
        t2 <- t + Y2
      }
      if(n3 == 1){
        Y3 <- rgamma(1, 1, 10)
        t3 <- t + Y3
      }
    }
    #case 4: [store open] someone leaves server 3 before someone arrive at server 3 
    #and before someone arrives at server 2 and before someone arrives at server 1
    else if(t3 == min(t1, t2, t3, Ta) & t3 <= T3){
      t <- t3
      Nd <- Nd + 1
      n3 <- n3 -1
      D3[Nd] <- t
      if(n3 == 0){
        t3 <- 9999
      }
      else{
        Y3 <- rgamma(1, 1, 10)
        t3 <- t + Y3
      }
    }
    if(min(Ta, t1, t2, t3) >= T3) break
  }
  #case 5: [store closed] someone still in store
  if((min(Ta, t1, t2, t3) >= T3) & (n1 > 0 | n2 > 0 | n3 > 0)){
    while(n1 > 0 | n2 > 0 | n3 > 0){
      if((t1 == min(t1, t2, t3)) & n1 > 0){
        t <- t1
        n1 <- n1 - 1
        n2 <- n2 + 1
        D1[Na - n1] <- t
        if(n1 > 0){
          Y1 <- rgamma(1, 1, 10)
          t1 <- t + Y1
        }else{
          t1 <- 9999
        }
        if(n2 == 1){
          Y2 <- rgamma(1, 1, 10)
          t2 <- t + Y2
        }
      }
      else if((t2 == min(t1, t2, t3)) & n2 > 0){
        t <- t2
        n2 <- n2 - 1
        n3 <- n3 + 1
        D2[Na - n1 - n2] <- t
        if(n2 > 0){
          Y2 <- rgamma(1, 1, 10)
          t2 <- t + Y2
        }else{
          t2 <- 9999
        }
        if(n3 == 1){
          Y3 <- rgamma(1, 1, 10)
          t3 <- t + Y3
        }
      }
      else if((t3 == min(t1, t2, t3)) & n3 > 0){
        t<- t3
        n3 <- n3 - 1
        Nd <- Nd + 1
        D3[Nd] <- t
        if(n3 > 0){
          Y3 <- rgamma(1, 1, 10)
          t3 <- t + Y3
        }else{
          t3 <- 9999
        }
      }
    }
  }
  
  cust.wait[i] <- mean(D3-A)
  i <- i + 1
}

mean(cust.wait)
```

**(3) grocery store**
```{r}
#three tellers with their own lines = lambda/3
N <- 100
cust.wait <- numeric(N)
Tp <- numeric(N)
depart <- numeric(N)
Timepast <- numeric(N)
i <- 0
lambda.rate <- function(t){
  if(t >= 4 & t <= 5){
    lambda <- 6/3
  }else{
    lambda <- 4/3
  }
  return(lambda)
}

for (i in 1:N) {		#starting the loop and initializing
  t<-0				#time starts at 0
  T<-8				#time is 8 hours
  A <- 0
  D <- 0
  Na <-0
  Nd <-0
  u1 <-runif(1)
  n  <-0
  Td <- 9999
  
  while(t<T){
    case <-0
    while (case==0){
      u2 <-runif(1)
      lt <- 1/ ((t -4)^2 + .5)
      lambda <- lambda.rate(t)
      lambda.t <- lt/lambda
      if (u2 <= lambda.t)
      { u1 <- runif(1)
      T0 <- t  - 1/lambda*log(u1)
      Ta <- T0
      case<- case + 1
      }
    }
    
    #Case 1
    
    if (Ta <= Td & Ta <= T){
      t <- Ta				#updating time variable
      Na <- Na + 1			#add additional arrival at time ta
      n <- n+1			#add additional customer
      A[Na] <-t			#output data (customer Na arrived at time t
      
      case <- 0
      while (case == 0){
        u2 <- runif(1)
        lt <- 1/ ((t -4)^2 + .5)
        lambda <- lambda.rate(t)
        if (u2 <=  (lt/ lambda)){ 
          u1 <- runif(1)
          Tt <-  t  - (1/lambda)*log(u1)   #generate time of next arrival
          Ta <- Tt			 #update time	
          case<- case + 1
        }
      }
      if (n==1){
        Y <- rgamma(1,1,10)
        Td <- t + Y		
      }
    }
    
    
    #CASE 2
    
    if (Td < Ta & Td <= T){
      t <- Td
      n <- n-1
      Nd <- Nd + 1
      D[Nd] <- t
      if (n == 0) {
        Td <- 99999 
      }
      else if (n > 0){	
        Y <-  rgamma(1,1,10)
        Td <-t + Y
      }
    }
    
    
    
    
    
    #Case 3
    
    if (min(Ta, Td) > T & n > 0){
      while (n > 0){
        t <- Td
        n <- n - 1
        Nd <- Nd + 1
        D[Nd] <-  t
        if (n > 0){
          Y <- rgamma(1,1,10)
          Td <- t + Y
        }
      }
      
    }
    
    #Case 4
    
    if (min(Ta, Td) > T & n == 0){
      Tp <- max(t - T, 0)
      Timepast[i] <- Tp
    }
  }
  
  cust.wait[i] <- sum(D - A)/Na
  depart[i] <-t
  i <- i+1
  
  
}
mean(cust.wait)
```


